### 2017-03-05 eda_report.rnw
Delivered EDA report to Guido.

\documentclass[12pt]{article}

\usepackage{algorithm2e} 
\usepackage{amsmath} \usepackage{amsthm} 
\usepackage{amsfonts} 
\usepackage{bbm} \usepackage{color,soul} 
\usepackage{framed} 
\usepackage[margin=0.5in]{geometry} \usepackage{hyperref} 
\usepackage{mathtools} 
\usepackage[dvipsnames]{xcolor}  
\newtheorem{theorem}{Theorem}[section] 
\newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  
\newcommand{\D}{\mathrm{d}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil} \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\SetKwInput{KwInput}{Input} 
\SetKwInput{KwOutput}{Output}

\begin{document}

\title{Exploring the ``Party City'' effect in the NBA}
\author{Andreas Santucci, Eric Lax}
\maketitle

\subsection*{Motivation}
It's not unreasonable to think that game performance can be affected (in part) by what takes place off the basketball court. We are interested to see if teams exhibit a decline in performance the day following a game in a party city. We exploit data on bookmaker spreads: the expected score differential between two teams after conditioning on observable performance. We expect a team to meet the spread half the time; if not, the bookmakers would open up an opportunity for arbitrage over the long-term. We construct a model which attempts to estimate the causal effect of  visiting a ``party-city'' the day before a game on the probability of meeting the spread.

\paragraph{Defining a party city}
We define a \emph{party city} to be one of the following: Atlanta, Brooklyn, Los Angeles, or Miami. We then look for other teams (not from one of these cities)
who travel to a party city and then play a game in another city the day after; each of the thirty NBA teams encounters this situation on average twice per season.

\subsection*{A simple first model}
What if we pick up the effect of a busy play-schedule, or travel from a far-away city? We consider comparing the difference in performance when a team has visited a party city the day before to the difference in performance when a team has visited any other city the day before. We estimate the probability of meeting the spread
using a logistic regression. Model sketch:
$$
\texttt{meet spread} \sim \substack{\texttt{days since} \\ \texttt{last game}} + \substack{\texttt{games played} \\ \texttt{last week}} + \substack{\texttt{team} \\ \texttt{ranking}} + \substack{\texttt{time}\\ \texttt{difference}} + \substack{\texttt{distance} \\ \texttt{traveled}} + \texttt{party}
$$

\subsection*{Challenges} Each of the thirty NBA teams runs into our quasi-experiment only a few times
per season, so our effective sample size relatively small. One possible remedy is to consider a more granular model, analyzing individual player performance. Another challenge is that our model relies
on spreads as input, for which we would like to know which factors are taken into account, especially
in so far as this model may change from season to season. Although
we cannot directly observe how the spreads model is built, we can observe how well different features
predict the probability of meeting the spread, and from this back out or infer what one such model 
may be.

\subsection*{References}
``The Benefits of College Athletic Success: An Application of the Propensity Score Design'', by Michael L. Anderson. March, 2017. \href{http://www.mitpressjournals.org/doi/pdfplus/10.1162/REST_a_00589}{paper}

\newpage

\section*{Exploratory Analysis}
Here, we explore a few aspects of our data, which are collected from
\href{http://www.goldsheet.com/histnba.php}{gold sheet}. We utilize 2008-2012 data. We start with a sanity check: if spreads are set correctly, they should be met on average half the time; otherwise an arbitrage opportunity exists. This
is verified.

<<echo=F,results='hide',include=FALSE>>=
require(data.table)
require(ggplot2)
require(magrittr)

load(file = 'tmp_data/spreads_wdist.RData')
load(file = 'tmp_data/standings.RData')

data[, season := gsub('_20', '_', season)]

standings <- subset(standings, select = c('team', 'season', 'pct'))
standings[, pct := shift(pct, type = 'lag'), by = team]

data <- merge(data, standings, by = c('season', 'team'), all.x = T)
data <- data[data[, head(.I, 82), by = list(season, team)]$V1]

### Define a Party Variable
party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
party.rgx <- paste0('(', party.cities, ')', collapse = "|")

data[, party := 0]
data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]

data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
@

<<echo=F,out.height="6in",warning=F>>=

### How often are teams meeting the spread?
hist(data[, mean(outcome == 'W'), by = list(season, team)][, V1],
     breaks = seq(0.3, 0.75, by = 0.025),
     main = "How often are teams meeting the spread?",
     xlab = "Proportion of games for which a team-year meets the spread")
@

We also check how many observations we will have to test our hypothesis. On average, each team
only encounters our quasi-experiment twice per season. This motivates the need for a more granular
model, which perhaps might look at the party effect on individual performance. We could
further break this down into offensive performance (shots scored per minute of playtime)
and defensive performance, perhaps.

<<echo=F,out.height="5.5in">>=
#data[, .N, by = list(season, team)]
data[party==1, .N, by = list(season, team)][, N] %>%
    hist(main = "Number of party days by season-team", xlab = "Number of days", ylab = "Frequency")
@

\subsection*{Arguing which factors are taken into account in the spread model}
We argue that the input spreads data has already taken into account (at least some of)
a team's ability, in so far as ability can be proxied by last season's average win-rate.

<<echo=F,warning=F,fig.pos='H'>>=

### TODO : Look at outliers here!!
### TODO : Remove playoff games.
m <- data[,#grepl(party.rgx, game.loc) & !grepl(party.rgx, team),
          list(meet.spread = mean(outcome == 'W'), pct = unique(pct), .N), by = list(team, season)]
ggplot(m, aes(x = pct, y = meet.spread)) +
    geom_point() +
    facet_wrap(facets = ~season) + 
    labs(title = "In select seasons, lagged performance not predictive of meet the spread",
         x = "Percentage of games won last season",
         y = "Proportion of games for which spread was met in current season")
@

We remark that in seasons starting in 2008-2010, there is almost no correlation between 
performance in the prior season and proportion of games for which spread was met in the current season. In 2011-12 and 2012-13 we have positive correlation between these two variables, which is interesting; we do not expect this. It is indicative
of a change in the way spreads are calculated, and is worthy of further investigation.

We now also argue that clearly the number of days since last game has been factored into spread data.

<<>>=
m[, cor(meet.spread, pct, use = 'complete'), by = season]
@

We see that on average, teams are meeting the spread half the time no matter if they've played
their most recent game one day ago or one week ago. There are a few outliers when teams are allowed
over one week rest; we note that these games are not of interest to our analysis, and may be excluded
without problem.

<<echo=F,out.height="6in",warning=F,fig.pos='H'>>=
m <- data[, list(meet.spread = mean(outcome == 'W'), .N), by = ndays.lgame][order(ndays.lgame)]
ggplot(m, aes(x = ndays.lgame, y = meet.spread, label = N)) +
    geom_point() +
    geom_label(nudge_y = -.04) +
    geom_hline(yintercept = 0.5) + 
    labs(x = "Number of days since last game", y = "Met the Spread",
         title = paste("Proportion of games which meet spread",
                       "as a function of rest-time",
                       "(Counts indicate number of observations used)", sep = "\n"))
@

The same is true for distance traveled as well.

<<echo=F,warning=F>>=
m <- data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(season, travel.dist = round(travel.dist, digits = -2))]
ggplot(m, aes(x = travel.dist, y = meet.spread, label = N)) +
    geom_point() +
    geom_label(nudge_y = -.01) +
    geom_hline(yintercept = 0.5) +
#    facet_wrap(facets = ~season) + 
    labs(x = "Travel Distance", y = "Met the Spread",
         title = paste("Proportion of games which meet spread",
                       "as a function of travel distance",
                       "(Counts indicate number of observations used)", sep = "\n"))

@

<<>>= 
m[, cor(travel.dist, meet.spread, use = 'complete')]
@

When we partition the above plot by season, we see that there is some linear correlation
between distance traveled and proportion of games for which spread was met; however, this correlation
heavily influenced by a few outliers (in particular, these outliers' averages are calculated based on very few observations).

<<>>=
m[, cor(travel.dist, meet.spread, use = 'complete'), by = season]
@

\section*{A tabular look at the effect of partying by team}

Lastly, we take a look at the effect of partying by season-team. Notice that
some teams are more effected than others; we hypothesize that demographics of the 
team influence the effect size. In particular, younger teams with unmarried players may
exhibit larger variability in performance after visiting party cities.

Some notable examples include the Denver Nuggets, and the Philadelphia 76'ers.
Also of importance: the San Antonio Spurs consistently have the oldest team, no matter which
statistic is used, and they also have a coach which is notoriously strict on partying; for this team, unsurprisingly, the party effect moves in the opposite direction. For 2008-2012 historical data, we display below the proportion of games for which
each team meets the spread under treatment as compared to without, along with the number of observations used to calculate the statistic.

<<>>=

data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(team, party)][order(team, party)]
@

We might also consider that the effect sizes may change as a function of how competitive the team is. Perhaps we may consider running a separate analysis for teams in the bottom quantile of last-year's rankings.
We also remark that there is not a lot of data supporting each mean estimate for
the treatment effect by team (especially at the season-team level), and this does pose issues.
\subsection*{A first model}
For the 2008-2012 seasons, we fit the following model.

<<>>=
load(file = 'tmp_data/spreads_with_standings.RData')
model <- glm(outcome == 'W' ~ party + weekend + ndays.lgame + pct +
                 travel.dist + team + season + team:season,
             family = 'binomial', data = data)

summary(model)$coef[c(1:6, sample(7:nrow(coef(summary(model))), size = 6)), ]
@

\section*{Problems, Extensions}

We remark that the magnitude and sign of the coefficient on \texttt{party} are of practical
importance. Although we don't achieve statistical significance, we may be able to come up with
a formulaic way to determine which teams are susceptible to the treatment effect, e.g.
teams with mostly young and unmarried demographic, and run a separate analysis for this group. It's also noteworthy that the model was teetering on weak statistical significance
before adding in 2008-2009 data, and the results could change slightly as we include more historical
data (either more recent or more past).

\end{document}