### 2017-05-06 poster/poster.tex
Init commit. Nice templated outline.

\documentclass[24pt, a0paper, portrait]{tikzposter} 
 \usepackage{algorithm2e} \usepackage{amsmath} \usepackage{amsthm} \usepackage{amsfonts} \usepackage{bbm} \usepackage{color,soul} \usepackage{framed} \usepackage{hyperref} \usepackage{mathtools} \usepackage[utf8]{inputenc} 
 \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-light on subsequent day sports performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle

\block{~} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court. We are interested to see if teams exhibit a decline in performance the day following a game in a party city. We exploit data on bookmaker spreads, the expected score differential between two teams after conditioning on observable performance. We expect a team to meet the spread half the time, since this is how bookmakers minimize risk on their end. We construct a model which attempts to estimate the causal effect of  visiting a "party-city" the day before a game on the probability of meeting the spread. }

\begin{columns}     \column{0.4}     \block{Data Sources}{
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
    }       \column{0.6}     \block{NBA model}{Here, \vspace{4cm}} \end{columns}

\begin{columns}     \column{0.5}     \block{Profits on MLB Betting in 2016}     {         \begin{tikzfigure}             \includegraphics[width=0.4\textwidth]{figures/mlb_bets}         \end{tikzfigure}     }     \column{0.5}     \block{Description of the figure}{
      On the left, we plot profits as a function of time over the 2016 season. For each relevant
      game, we place a bet if our model instructs us of odds sufficiently different from what
      the betting house suggests. We account for the bookmaker's cut. In the early season,
      at its worst we are out of pocket \$750 dollars. By season end, we have recouped all of
      our initial investment plus an additional \$13.5k.
    }
    \note[         
      targetoffsetx=-8cm,          
      targetoffsety=-6.5cm,          
      width=0.5\linewidth         
    ]{e-mail: \texttt{santucci@stanford.edu}}   \end{columns}   \end{document}